---
title: "R Notebook"
output: html_notebook
---

Question 1
Load in the data, and have a look at it. How many variables does it have? How many observations? What are the variable names? What type of variables does it have?

```{r}
#load in required libraries
library(tidyverse)
library(janitor)
```

```{r}
#load in data
income <- read_csv("data/state_income_data.csv") %>% clean_names()
#view(income)
```

```{r}
#number of rows and columns
dim(income)
```
```{r}
#name of columns
names(income)
```

```{r}
#column types
glimpse(income)
```

Question 2
Select the columns which contain the information from years 2002 to 2008. Ensure you keep the Index and State variables as well.

```{r}
income %>% 
select(index:y2008)
```


Question 3
Rename the Index column to first_letter. Hint: look up the rename() function.

```{r}
income <- income %>% 
  rename(first_letter = index)
```


Question 4
Find the seven states which start with an A,B, or C.
```{r}

states_A_B_C <- income %>% 
  filter(first_letter %in% "A" | first_letter %in% "B" |  first_letter %in% "C")
```


Question 5
Find which of these seven states have the greatest income in 2008.

```{r}
states_A_B_C  %>% 
  select (state, y2008) %>% 
  slice_max(y2008)
```



Question 6
Calculate the change in income for each state of these seven states (states starting with A,B, or C) between 2002 and 2008. Save the income change in a new column called income_change.

```{r}
states_A_B_C_inc <- states_A_B_C %>% 
  mutate(income_change = y2008 - y2002)
```


Question 7
Find which of these seven states income change is greater than 500,000 (i.e. an increase of 500,000 or decrease of 500,000).
```{r}
states_A_B_C_inc %>% 
  filter(abs(income_change) > 500000)
```

Question 8
Calculate the mean and median income change for all seven states.

```{r}
states_A_B_C_inc %>% 
  summarise(mean = mean(income_change), median = median(income_change) ) #%>% 
  #summarise(median(income_change))
```


Question 9
Go back to the original dataset which contains all states. Write a pipe which selects the State and Y2008 variables, arranges the income from highest to lowest, filters all those states which have a income of greater than 1 million, and then finally chooses the top 10 wealthiest states.

```{r}
income %>% 
  select(state, y2008) %>% 
  arrange(desc(y2008)) %>% 
  filter(y2008 > 1000000) %>% 
  head(10)
```

