---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
```


1
Load in the country_data.csv file and take a look at it. What needs changed in this to make it tidy? Which columns need made into long format, and which need spread to rows?

```{r}
country <- read_csv("data/country_data.csv") %>% clean_names()
```
```{r}
view(country)
```


2
Do the appropriate pivoting of your columns using R.


```{r}
#create one year column, store values in temp column called values
country_longer <- country %>%
 pivot_longer(cols = starts_with("y"), 
               names_to = "year", 
               values_to = "values")

view(country_longer)

country_longer_wider <- country_longer %>%
pivot_wider(names_from = type, 
              values_from = values)

#view(country_longer_wider)
```


3
Separate the country variable so you have the name in one column and the number in another.

```{r}
country_split <- country_longer_wider %>% 
  separate(country, c("number", "name"), sep = "\\." )

#view(country_split)
```


4
Unite the code_letter and code_numb variables so that they are both in one column, with no spaces separating them.

```{r}
country_final <- country_split %>% 
  unite(code, c("code_letter", "code_numb"), sep ="")

#view(country_final)
```

