---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
```


Question 1.

Read in the data and investigate it, e.g. variables, variable types, dimensions etc.


```{r}
wine <- read_csv("wine_ratings.csv") %>% clean_names()

dim(wine)
glimpse(wine)
#view(wine)

```


Question 2.

Select columns title, taster_name, points, price, and country, then drop all NAs. Save into a variable called wine_subset. Use this variable for all subsequent tasks.

```{r}
wine_subset <- wine %>% 
  select(title, taster_name, points, price, country) %>% 
  drop_na()

#view(wine_subset)
```


Question 3.

Find the 5 countries with the highest mean points.
```{r}
wine_subset %>% 
  group_by(country) %>% 
  summarise(mean_points = mean(points)) %>% 
  arrange(desc(mean_points)) %>% 
  head(5)
```


Question 4.

Find the 5 countries with the lowest mean prices.

```{r}
wine_subset %>% 
  group_by(country) %>% 
  summarise(mean_prices = mean(price)) %>% 
  arrange(mean_prices) %>% 
  head(5)
```


Question 5.

Does the result differ if you use the median instead of the mean for Questions 3 and 4?

```{r}
wine_subset %>% 
  group_by(country) %>% 
  summarise(median_points = median(points)) %>% 
  arrange(desc(median_points)) %>% 
  head(5)

wine_subset %>% 
  group_by(country) %>% 
  summarise(median_prices = median(price)) %>% 
  arrange(median_prices) %>% 
  head(5)

```


Question 6.

Find the names of all the tasters using distinct. Then take a random sample of 5 names using sample_n. Save these 5 names into a new variable called taster_sample which should be a character vector containing only the names.

```{r}
taster_sample <- wine_subset %>% 
  distinct(taster_name) %>% 
  sample_n(5) %>% 
  pull()

#check if this is a character vector
class(taster_sample)
```


[Hint: You can use the function pull to extract a single column as a vector.]

Question 7

Do the same thing for country, i.e. create a character vector called country_sample using distinct and pull, but this time use sample_frac to pick 10 countries at random.

[Hint: What is the proportion of countries you want? \(\frac{Part}{Whole}\)]

```{r}
(country_sample <- wine_subset %>% 
  distinct(country) %>% 
  sample_frac(0.3) %>% 
  pull())

#check if this is a character vector
class(country_sample)
```


Question 8

Filter the rows of wine_subset so you are left only with data from the 5 tasters in your taster_sample and from the countries in your country_sample.

```{r}
wine_subset %>%
  filter(country %in% country_sample & taster_name %in% taster_sample)
  
```


Question 9

Using your wine_sample, create a new column difference_from_avg which contains the difference between each wine’s rating (points) and the average rating for that taster, then arrange the differences to see the highest positive difference first. [Hint: you will first have to calculate the average points for each taster]

```{r}
#find the average rating for each taster and store in a dataframe
average_taster_rating <- wine_subset %>% 
  group_by(taster_name) %>% 
  summarise(average_rating = mean(points))

#join our initial DF with the newly created DF for avg. taster ratings, using taster name to join      
wine_subset <- inner_join(wine_subset, average_taster_rating, by = "taster_name" )

#add new column to record calculated difference in rating for each wine compared to average rating of the taster
wine_subset <- wine_subset %>% 
mutate(difference_from_avg = round(points - average_rating)) %>% 
  #use absolute values to ignore sign
  arrange(desc(abs(difference_from_avg)))

wine_subset
```

